---
title: "What Size Houses Should They Build?"
author: "Mr. Boerman"
format: html
editor: visual
execute:
  warning: false
  messages: false
---

# Load Packages
```{r}
#| label: setup

library(tidyverse)
library(here)
library(readxl)
library(scales)
library(knitr)
```


# Read Data - Pre-Cleaned
```{r}
# This is a file that I downloaded from the link above,
# opened it in Excel,
# deleted the unneeded rows at the top,
# renamed the columns to be code-friendly (lowercase and _),
# ensured the columns contained one type of data only,
# saved as a csv file,
# and uploaded to my posit working folder. 

# This approcah is less reproducible, but easier.

cleaned_data <- read_csv(
  file = here("input_data/household_size_cleaned.csv")
)
```


```{r}
formatted_data <- cleaned_data |> 
  filter(year == 2025) |> 
  select(!c(total_count, average)) |> 
  pivot_longer(
    cols = c(!year),
    names_to = "household_size",
    values_to = "number_of_households"
) |> 
  mutate(household_size = case_when(
    household_size == "hhsize_1" ~ 1,
    household_size == "hhsize_2" ~ 2,
    household_size == "hhsize_3" ~ 3,
    household_size == "hhsize_4" ~ 4,
    household_size == "hhsize_5" ~ 5,
    household_size == "hhsize_6" ~ 6,
    household_size == "hhsize_7+" ~ 7
  )
) |> 
  arrange(household_size)
```

# Explore Data
```{r}
#| label: plot-histogram
#| fig-width: 8
#| fig-height: 5

formatted_data |> 
ggplot(
  aes(x = household_size, y = number_of_households)) +
  geom_col(fill = "steelblue", color = "black", width = 0.7) +
  labs(
    title = "Household Size Distribution (2025)",
    subtitle = "Number of total households by size",
    x = "Household Size",
    y = "Number of Households"
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = c(1:7)
  ) +
  scale_y_continuous(
    expand = c(0,0)
  ) +
  theme_minimal(base_size = 12)

```

# Create Variables of Interest
```{r}
data_with_percents <- formatted_data |> 
  mutate(percent_of_total = number_of_households / sum(number_of_households)) |> 
  mutate(cumulative_percent = cumsum(percent_of_total))
```

```{r}
data_with_percents |> 
  ggplot(aes(
    x = household_size,
    y = cumulative_percent
  )) +
  geom_step() +
  geom_point() +
  scale_x_continuous(
    breaks = 1:nrow(data_with_percents),
    labels = data_with_percents$household_size,
    # expand = c(0,0)
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 1),
    # expand = c(0, 0)
  ) +
  labs(
    title = "Cumulative Distribution of Household Sizes (2025)",
    subtitle = "Share of Households Accounted for at Each Size",
    y = "Percent Accounted For",
    x = "Household Size"
  ) +
  theme_minimal()
```


# State Question
I think that the average household size is not the most common household size. 
I think that there are many small households (1-2 people) who needs small houses (1-2 bedrooms) and a decent number of large households (5-7 people) who needs large houses (5-7 rooms). 
By producing houses with 3 beds, the company is cutting out the ends of the spectrum from their customer base. 

# Answer Question
To prove this, I will need to show the number of households that would buy a house that is 1 bed, 2 bed, 3 bed, etc.

I can use the current variable of "percent of total" as a proxy, assuming that the number of bedrooms equals the household size (mom and dad may share a room, but buy 1 extra for growign family, and office, playroom, etc.)

If the percent of total for any given household is large, the company should make a bunch of houses with that many rooms.

In fact, the percent of total of household size should equate to the percent of total of houses built by the company!

If the percent of total matches what they make now, they are in good shape.
If not, they need to change what size houses they build. 

```{r}
data_with_percents |> 
  mutate(
    percent_formatted = paste0(round(percent_of_total * 100, 0), "%")
  ) |> 
  select(household_size, percent_formatted) |> 
  kable()
```



# Draw Conclusion
...