---
title: "Car Gas Plots"
author: "Mr. Boerman"
format: html
editor: visual
execute:
  warning: false
---

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
library(tidyr)

```

```{r}
# Read the data directly from GitHub
gas_log <- read_csv("https://raw.githubusercontent.com/michaelboerman/bmeg_226/main/input_data/car_gas_log.csv") |> 
  mutate(
    date = mdy(date),
    mpg = miles / gallons,
    cost_per_gallon = cost / gallons,
    cost_per_mile = cost / miles,
    year = year(date),
    month = month(date, label = TRUE)
  )

# Display the processed data
head(gas_log)
```


# Question 1: Is This Data Normally Distributed?

### Step a: Look at the data
```{r}
gas_log |> 
  ggplot(aes(x = mpg)) +
  geom_histogram(
    binwidth = 15,
    color = "white",
    fill = "black"
  ) +
  labs(
    title = "Distribution of MPG per tank",
    x = "Miles per Tank",
    y = "Number of Instances"
  ) +
  theme_minimal()
```
Maybe kinda normal with a skew. 

Experiment with how the chart changes when we change the bin width. 
Some bin widths make me think there's no way it's normal.
Other bin widths lead me to believe it could be normal.

Also, the x and y intervals don't make much sense.


```{r}
gas_log |> 
  select(mpg) |> 
  pull() |> 
  shapiro.test()
```

What does this p-value mean?
How do you find out?

```{r}
gas_log |> 
  select(mpg) |> 
  pull() |> 
  qqnorm()
```

```{r}
gas_log |> 
  select(mpg) |> 
  pull() |> 
  log() |> 
  shapiro.test()
```

```{r}
gas_log |> 
  select(mpg) |> 
  pull() |> 
  log() |> 
  qqnorm()
```


# Question 2: Does My Mileage Per Tank Change Over Time?
Goal: Evaluate whether there is a staistically significant difference in MPG per year. 
```{r}
annual_summary <- gas_log |>
  group_by(year) |>
  summarise(
    avg_mpg = mean(mpg),
    avg_cost_per_gallon = mean(cost_per_gallon),
    total_miles = sum(miles),
    total_cost = sum(cost),
    n_fillups = n()
  )

```

```{r}
gas_log |>
  ggplot(aes(x = month, y = mpg)) +
  geom_boxplot(fill = "lightblue", alpha = 0.6) +
  geom_jitter(width = 0.1, alpha = 0.5, color = "navy") +
  labs(
    title = "Monthly Variation in Fuel Efficiency",
    subtitle = "Boxplots show distribution; points show individual fill-ups",
    x = "Month",
    y = "MPG"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
gas_log |>
  group_by(year) |>
  summarise(
    `Total Miles` = sum(miles),
    `Total Cost ($)` = round(sum(cost), 2),
    `Avg MPG` = round(mean(mpg), 1),
    `Avg $/Gallon` = round(mean(cost_per_gallon), 2),
    `Fill-ups` = n()
  ) |>
  knitr::kable()

```
